---
title: "individual2.Rmd"
author: Samuel Cole
date: 4/18/2019
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
names(acc)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
names(acc)
```
**Part one: Accident data**
Are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?
```{r}
acc$DAY_WEEK
library(ggplot2)
ggplot(data = acc, aes(x=DAY_WEEK)) + geom_histogram(color="black", fill="light blue") + ylab("Total Accidents") + xlab("Day of the week") + ggtitle("Total Accidents on each Given week day")
```
What time of the day do accidents happen (use variable HOUR)?
```{r}
acc[acc$HOUR == 99] <- NA
#We have to change 99 to null in order to graph this as a per hour accident graph. 99 is the NULL hour, because the 0 hour is from midnight to 1.
ggplot(data = acc, aes(x=HOUR)) + geom_histogram(color="black", fill="light blue") + ylab("Total Accidents") + xlab("Hour of the Day") + ggtitle("Total Accidents Per Hour on a Given Day")
```
What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?
```{r}


```

**Part two: Connecting data**

Load the person table. Identify drivers (PER_TYP == 1, see fars manual) and subset on them.
```{r}


```
Join accident and driver table (work out which variable(s) to use)
```{r}


```
Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!
```{r}


```

**Part three: Connecting data and make maps**

Now plot a choropleth map of accidents on a county level
```{r}


```
Read the description for the STATE and COUNTY columns in the FARS manual.
The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication
Use readxl::read_xlsx to read in the GLCs
```{r}


```
Visualize the number of accidents in each county using a choropleth map. To do this, you need to join the accident data, GLCs, and appropriate map data.
```{r}


```
Can you find seasonal trends in accident occurrences in different states? Is summer or winter more dangerous? Why?
```{r}


```
