---
title: "individual2.Rmd"
author: Samuel Cole
date: 4/18/2019
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
names(acc)

```
**Part one: Accident data**
Are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?
```{r}
library(ggplot2)
ggplot(data = acc, aes(x=DAY_WEEK)) + geom_histogram(color="black", fill="light blue") + ylab("Total Number of Accidents") + xlab("Day of the week") + ggtitle("Total Accidents on each Given Week day") + scale_x_continuous(breaks= c(1, 2, 3, 4, 5, 6, 7), labels = c("Sunday", "Monday", "Tuesday","Wednesday", "Thursday", "Friday", "Saturday"))
```
There are more accidents on Friday, Saturday, and Sunday. These are all considered the weekend, so the weekend works as well. 
What time of the day do accidents happen (use variable HOUR)?
```{r}
acc$HOUR[acc$HOUR == 99] <- NA
#We have to change 99 to null in order to graph this as a per hour accident graph. 99 is the NULL hour, because the 0 hour is from midnight to 1.
ggplot(data = acc, aes(x=HOUR)) + geom_histogram(color="Black", fill="light green") + ylab("Total Number of Accidents") + xlab("Hour of the Day") + ggtitle("Total Accidents Per Hour on a Given Day") + scale_x_continuous(breaks= c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23))
```
Most accidents are during rush hour around when people get off work. The graph starts to go up tremendously around the 15th hour of the day which is 3 in the afternoon, which is the start of when people start to get off work. 

What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?
```{r}
ggplot(data = acc, aes(x=DRUNK_DR)) + geom_histogram(color="black", fill="lightpink3") + ylab("Total Number of Accidents") + xlab("Number of Drunk Drivers involved in the Crash") + ggtitle("Number of Drunk Drivers involved in each Accident") + scale_x_continuous(breaks= c(0, 1, 2, 3)) + stat_bin(aes(y=..count.., label=..count..), geom="text", vjust=-.5) 
library(tidyverse)

drunkDriving <- acc %>% filter(DRUNK_DR > 0)
count(drunkDriving)

```
As we can see, there is 8769 cases where a drunk Driver was involved in the crash. The graph shows the overall layout of the number of Drunk Drivers per crash.

**Part two: Connecting data**

Load the person table. Identify drivers (PER_TYP == 1, see fars manual) and subset on them.
```{r}
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
names(acc)

drivers = per %>% filter(PER_TYP == 1)
```
Join accident and driver table (work out which variable(s) to use)
```{r}
acc.drivers <- inner_join(acc, drivers, by= "ST_CASE")
```

Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!
```{r}
acc.driver.clean <- acc.drivers %>% group_by(DAY_WEEK, HOUR.x, SEX)
tally(acc.driver.clean)
```

**Part three: Connecting data and make maps**

Now plot a choropleth map of accidents on a county level
```{r}
library(classdata)
library(tidyverse)
library("readxl")
CountyData <- read_excel("County Codes United States.xlsx")
county <- map_data("county")
ggplot(county, aes(x=long, y = lat)) + geom_polygon(aes(group=group))

```
Read the description for the STATE and COUNTY columns in the FARS manual.
The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication
Use readxl::read_xlsx to read in the GLCs
```{r}


```
Visualize the number of accidents in each county using a choropleth map. To do this, you need to join the accident data, GLCs, and appropriate map data.
```{r}


```
Can you find seasonal trends in accident occurrences in different states? Is summer or winter more dangerous? Why?
```{r}


```
